---
title: 'Final project: Anime Recommendation System with Rating Prediction'
author: "Yagna sri madepally"
date: "2024-04-30"
output:
  word_document: default
  pdf_document: default
---

### INTRODUCTION:

In this project, I aim to develop an anime recommendation system with rating prediction using two datasets: Anime and Rating. 

Source: https://www.kaggle.com/datasets/CooperUnion/anime-recommendations-database

Anime.csv: This dataset contains 7 columns and 12,294 rows. Each row represents an anime entry with the 
following attributes:

  •	anime_id: Unique identifier for each anime on myanimelist.net.
  •	name: Full name of the anime.
  •	genre: Comma-separated list of genres associated with the anime.
  •	type: Type of anime (e.g., movie, TV series, OVA, music, special, etc.).
  •	episodes: Number of episodes (1 if it's a movie).
  •	rating: Average rating of the anime, ranging from 1 to 10.
  •	members: Number of community members associated with the anime.

Rating.csv: This dataset consists of 3 columns and 46,986 rows. It represents user ratings for various anime, including:

  •	user_id: Randomly generated unique user identifier.
  •	anime_id: The ID of the anime that the user has rated.
  •	rating: The user's rating for the anime on a scale of 1 to 10.(-1 if watched but not rated)

The objective of this project is to build a recommendation system that can suggest anime titles to users based on their preferences.

Example of inspiration: Netflix Recommendation System

A similar recommendation system can be found on platforms like Netflix, where users are provided with personalized suggestions based on their viewing history and preferences. For instance, Netflix's "Users also watched" feature recommends titles that are similar to the ones users have previously watched or rated positively. This enhances the user experience by offering relevant content tailored to individual tastes.



## Step 1: Data Preparation/collection

Loading the required packages

```{r}
library(tidyverse)

library(e1071) #Using the e1071 package to calculate the skewness

library(treemap) #for creating a treemap

library(rsample)   # for resampling procedures

library(caret)  # for fitting KNN models

library(recipes)    # for feature engineering

library(tidymodels) #for tidy and unified modeling workflow

library(reshape2) #for reshaping and restructuring data

library(Matrix) #Matrix library for sparse and dense matrix classes and methods

library(stringr) #for string manipulation functions

library(class) #for various classification methods including k-Nearest Neighbors (kNN)

library(knitr)

library(DT)
```

Reading the data

```{r}
library(readr)
Anime<- read.csv('C:/Users/yagna/Desktop/MSIS/581/FINAL_PROJECT/anime.csv')
Rating<-read.csv('C:/Users/yagna/Desktop/MSIS/581/FINAL_PROJECT/ratings.csv')

```


Exploring the Anime and Rating datasets

```{r}
dim(Anime)
dim(Rating)

```

Glimpse of anime dataset

```{r}
glimpse(Anime)

```


Glimpse of rating dataset

```{r}
glimpse(Rating)

```

summary of anime and rating data sets

```{r}

summary(Anime)

summary(Rating)
```

## STEP 2: DATA CLEANING

Let's remove the duplicate values in Anime dataset and compare the no of rows:

```{r}
Anime_dup=unique(Anime)
dim(Anime_dup)  

```

we can see that the no of rows are: 12,294 which means we have zero duplicate values

Let's check for duplicate values in Rating dataset:

```{r}
Rating_dup=unique(Rating)
dim(Rating_dup)
```

It has zero duplicate entries among 46,986 entries.

Identifying the null values in Anime and Rating datasets: 

```{r}
Anime_null_values <-colSums(is.na(Anime_dup))
Rating_null_values <-colSums(is.na(Rating_dup))
```

Transposed view of Anime dataset null values:

```{r}
transposed_data <- t(Anime_null_values)

kable(transposed_data)
```

Transposed view of Rating dataset null values:

```{r}
transposed_data <-t(Rating_null_values)

kable(transposed_data)
```

Ratings dataset has zero null values.

The number of null values in Anime_dup dataset is relatively small compared to the overall dataset size and dropping them doesn't significantly affect the analysis

Dropping the null values in Anime:

```{r}
Anime_df <- Anime_dup[complete.cases(Anime_dup), ] 
```


Lets check for null values again:

```{r}

Anime_null_values <-colSums(is.na(Anime_df))

transposed_data<-t(Anime_null_values)


kable(transposed_data)
```

The null values are zero in Anime_df dataset

Now let us merge the datasets:

```{r}

merged_data <- merge(Anime_df, Rating_dup, by = "anime_id")

glimpse(merged_data)
```


Renaming the columns in merged_data:

```{r}

ColNames <- c("Anime_ID", "Name", "Genre", "Type", "Episodes", "Rating", "Members", "User_ID", "User_Rating")

colnames(merged_data) <- ColNames

glimpse(merged_data)
```


We have many special characters in the Anime Name. ex: Wolf&#039;s Rain
Lets remove the special characters from the Name

```{r}
# Defining patterns to remove
patterns <- c("&quot;", "\\.hack//", "&#039;", "A's", "I's", "&amp;")

# Iterate over patterns and remove them from the Name column
for (pattern in patterns) {
  merged_data$Name <- gsub(pattern, "", merged_data$Name)
}

```

Capitalize every starting letter in the Name column

```{r}
merged_data$Name <- str_to_title(merged_data$Name)

glimpse(merged_data$Name)
```

## STEP 3: Analysis

# 1. Anime Categories

```{r}

Anime_types<-merged_data %>%
  group_by(Type) %>%
  summarise(Total= n_distinct(Anime_ID)) %>%
  arrange(desc(Total))


# Calculate the percentage of each category
Anime_types <- Anime_types %>%
  mutate(Percentage = (Total / sum(Total)) * 100)
datatable(Anime_types)
```


```{r}
# Converting 'Type' variable to categorical using factor()
Anime_types$Type <- factor(Anime_types$Type)
str(Anime_types)

# Plotting a pie chart
ggplot(Anime_types, aes(x = "", y = Percentage, fill = Type, label = paste0(round(Percentage, 1), "%"))) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Anime Categories", fill = "Type", y = "Percentage") +
  geom_text(position = position_stack(vjust = 0.5), color = "black") +
  theme_void() +
  theme(legend.position = "right")
```

From the plot we can say that majority (43.4%) of anime is aired on TV.

# 2.Most Popular Anime: Community size vs. Ratings

```{r}
# Filtering the data to select the top 20 rows based on "Members"
top_20 <- merged_data %>%
  group_by(Name) %>%
  summarise(Total_Members = sum(Members),
            Rating = mean(Rating, na.rm = TRUE)) %>%
  top_n(20, Total_Members) %>%
  arrange(desc(Total_Members))

datatable(top_20)
```


```{r}
# Creating the bar plot of top 20 popular anime
ggplot(data = top_20,
       mapping = aes(x = reorder(Name, -Total_Members), y = Total_Members,fill = Rating)) +  # Reorder anime names based on Members
  geom_bar(stat = "identity") +
  geom_text(aes(label = Total_Members), vjust = 0.5, color = "black", size = 3, angle = 90) +  # Add labels inside the bar
  labs(x = "Anime Name", y = "Members") +
  scale_fill_gradient(low = "cornflowerblue", high = "darkblue")+
  ggtitle("Top20 Most Popular Anime: Community size vs. Ratings")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability


```

# 3.Exploring the different genres 

```{r}
glimpse(merged_data$Genre) 

```

We can see multiple genres are separated by a ",". now lets split the genre of each row seperated by a comma(,)

```{r}
# Extract distinct anime
distinct_genres <- merged_data %>%
  select(Anime_ID, Genre) %>%
  distinct()

glimpse(distinct_genres)
```



```{r}
# Split the string in each row to extract individual genres
genre_list <- strsplit(distinct_genres$Genre, ", ")

# Convert the list of lists into a single vector
all_genres <- unlist(genre_list)

# Counting the occurrences of each genre
genre_counts <- table(all_genres)

# Converting the result into a dataframe
genre_df <- as.data.frame(genre_counts)

# Renaming the columns
colnames(genre_df) <- c("Genre", "Count")

genre_df <- genre_df %>% 
  arrange(desc(Count))

glimpse(genre_df)
```

We have 43 different genre in our merged dataset

```{r}
# Creating the treemap plot

treemap(genre_df, 
        index = "Genre",  # Specify the index column
        vSize = "Count",  # Specify the column for the size of the tiles
        title = "Genre Distribution",  # Title of the treemap
        fontsize.title = 14,  # Font size for the title
        fontsize.labels = 10,  # Font size for the labels
        fontsize.labels.min = 6,  # Minimum font size for labels
        align.labels = list(c("center", "center")),  # Alignment of labels
        palette = "RdYlBu",  # Color palette for the tiles
        width = 10,  # Width of the plotting area
        height = 10  # Height of the plotting area
)
```

# 4: Finding potential outliers

Finding the distribution of the user rating to identify any potential outliers.

```{r}
# Sort merged_data by User_Rating in descending order
merged_data <- merged_data %>% 
  arrange(desc(User_Rating))

#generating a histogram of user rating 

ggplot() +
  geom_histogram(data = merged_data, aes(x = User_Rating), fill = "cornflowerblue", color = "black") +
  labs(title = "User Anime ratings distribution", x = "User rating", y = "Total")
```

We can see that -1 is a outlier in user ratings. -1(If the user has watched the anime but haven't rated it)

```{r}
skewness(merged_data$User_Rating)

```

The skewness is -0.96<0 proves that distribution is negatively skewed.

## STEP 4: Modeling

Before proceeding let us make sure to consider only valid scenarios.

Lets replace -1 (non rating) in User_rating column with the possible predicted rating.i.e,Let's predict user ratings for anime they haven't rated.

```{r}
#Identify unrated anime
unrated_anime <- merged_data %>%
  filter(User_Rating == -1)

glimpse(unrated_anime)
```

We can see that we have 10,426 unrated entries in our data frame.

```{r}
# Predict ratings for unrated anime
predicted_ratings <- unrated_anime %>%
  group_by(Name) %>%
  summarise(Predicted_Rating = mean(Rating))

# Merge predicted ratings with merged_data based on the anime name
merged_data <- merge(merged_data, predicted_ratings, by = "Name", all.x = TRUE)

# Replace "-1" with predicted ratings
merged_data$User_Rating[merged_data$User_Rating == -1] <- merged_data$Predicted_Rating[merged_data$User_Rating == -1]

# Remove the temporary column used for prediction
merged_data <- subset(merged_data, select = -c(Predicted_Rating))

merged_data <- merged_data %>% 
  arrange(desc(User_Rating))

# View updated dataset
summary(merged_data)
```

We have handled all the outliers. 

```{r}
dim(merged_data)
```

Merged_data includes columns: User_ID, Anime_ID, User_Rating

```{r}
colnames(merged_data)
```

splitting the dataset into training(70%) and testing(30%)

```{r}
merged_data$Liked <- ifelse(merged_data$User_Rating >= 7, 1, 0) #Creating Liked column where if rating is >7 then its values is 1 and if <1 then its value is 0
train_index <- createDataPartition(merged_data$Liked, p = 0.7, list = FALSE)
train_data <- merged_data[train_index, ]
test_data <- merged_data[-train_index, ]

# Preprocess the data
train_X <- train_data[, c("User_ID", "Anime_ID")]
train_y <- train_data$Liked
test_X <- test_data[, c("User_ID", "Anime_ID")]
test_y <- test_data$Liked

```

Training the kNN model using caret's train function

```{r}
model_knn <- train(x = train_X, y = train_y, method = "knn", trControl = trainControl(method = "cv", number = 5))

# Converting train_y to a factor with two levels
train_y <- as.factor(train_y) 
test_y <- as.factor(test_y)

# Make predictions on the test data
test_predictions <- predict(model_knn, newdata = test_X)

# Evaluate accuracy
accuracy <- mean(test_predictions == test_y)
cat("Accuracy:", accuracy, "\n")
```

This shows that our model is 83% accurate. It indicates that the kNN model performs reasonably well on the test data.

Anime recommendation function:

```{r}
recommend_similar_anime <- function(anime_id, n) {
  # Find the row corresponding to the input anime ID in the test data
  input_anime_row <- test_data[test_data$Anime_ID == anime_id, ]
  
  # Check if the input anime ID is present in the test data
  if (nrow(input_anime_row) == 0) {
    print("Anime not found in the dataset.")
    return(NULL)
  }
  
  # Extract features of the input anime
  input_features <- input_anime_row[, c("Genre", "Rating", "Members")]
  
  #Calculate similarity between input anime and all other anime based on features
  #lapply is a function in R used to apply a given function to each element of a list or vector, and returns a list containing the results. 
  
  anime_similarity <- lapply(unique(merged_data$Anime_ID), function(id) {
    anime_row <- merged_data[merged_data$Anime_ID == id, ]
    if (nrow(anime_row) > 0) {
      # Calculating the similarity score based on the intersection of genres
      similarity <- sum(input_features$Genre %in% anime_row$Genre) / length(input_features$Genre)
      return(c(id, similarity))
    } else {
      return(NULL)
    }
  })
  
  # Remove NULL values and convert to matrix
  anime_similarity <- matrix(unlist(anime_similarity), ncol = 2, byrow = TRUE)
  anime_similarity <- anime_similarity[complete.cases(anime_similarity), ]
  colnames(anime_similarity) <- c("Anime_ID", "Similarity")
  
  # Order by similarity and get top n similar anime
  similar_anime <- anime_similarity[order(-anime_similarity[, "Similarity"]), "Anime_ID"][1:n]
  
  return(similar_anime)
}
```

Function to map anime names to anime IDs:

```{r}
get_anime_id <-function(anime_name) {
  anime_id <- merged_data$Anime_ID[merged_data$Name == anime_name]
  if (length(anime_id) == 0) {
    print("Anime name not found in the dataset.")
    return(NULL)
  }
  return(anime_id)
}
```

Function to retrieve anime names

```{r}
get_anime_names <- function(anime_ids) {
  anime_names <- unique(merged_data$Name[merged_data$Anime_ID %in% anime_ids])
  return(anime_names)
}
```

Example:

```{r}
query_anime_name <- "death note"  
```


Replace with the anime name of the anime you want recommendations for.


```{r}
query_anime_name<-str_to_title(query_anime_name) #Handling case sensitive inputs.
query_anime_id<-get_anime_id(query_anime_name)
recommendations <- recommend_similar_anime(query_anime_id, 10) #Lets take 10 similar anime suggestions
recommended_names <- get_anime_names(recommendations)

cat(paste("Anime recommendations similar to", query_anime_name, "are:\n", paste(recommended_names, collapse = "\n"), "\n"))

```

